Probability Aggregation in Time-Series: Dynamic Hierarchical Modeling of Sparse Expert Beliefs

### 2013-09-29 paper.tex
Final commit before submit!

\begin{aug}
\author{\fnms{Ville A.} \snm{Satop\"a\"a}\corref{}\ead[label=e1]{satopaa@wharton.upenn.edu}},
\author{\fnms{Shane T.} \snm{Jensen}\ead[label=e3]{stjensen@wharton.upenn.edu}},
\author{\fnms{Barbara A.} \snm{Mellers}\ead[label=e4]{mellers@wharton.upenn.edu}},
\author{\fnms{Phil E.} \snm{Tetlock}\ead[label=e5]{tetlock@wharton.upenn.edu}},
\and
\author{\fnms{Lyle H.} \snm{Ungar}\ead[label=e2]{ungar@cis.upenn.edu}}

@@

This paper begins by describing our geopolitical database. The paper then introduces a dynamic hierarchical model for capturing the crowd belief. The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure. The model introduction is followed by the first evaluation section that uses synthetic data to study how accurately the two-step procedure can estimate parameter values. The second evaluation section applies the model to real-world forecasting data. The paper concludes with a discussion on future research directions and model limitations.

@@

To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}. The points have been jittered slightly to make overlaps visible. The darkness of the points is positively associated with the self-reported expertise. Given that the European bailout fund was ratified before November 1, 2011 and that the Nikkei 225 index finished trading at around 8,700 on September 30, 2011, the general trend of the probability forecasts tends to converge towards the correct answers. The individual experts, however, sometimes disagree strongly, with the disagreement persisting even near the closing dates of the questions.

@@

\subsection{Sampling Step}
\label{sampling_step}
Given that $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable. As a result, the parameter estimates tend to drift during the sampling process. A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.  In Section \ref{syntheticData} we provide a guideline for choosing the reference point. Denote the constrained version of the model by
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
X_{t, k}(1) &=& \gamma_k(1) X_{t-1, k}(1) + w_{t, k}\\
\boldsymbol{v}_{t, k} | \sigma^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k(1) \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}\left(0, \tau^2_k(1)\right),
\end{eqnarray*}
where the trailing input parameter emphasizes the constraint $b_{3} = 1$. Given that this version is identifiable, estimates of the model parameters can be obtained. Denote the estimates by placing a hat on the parameter symbol. For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively. These estimates are found via Gibbs sampling that only makes use of standard distributions. See Appendix \ref{appendix} for the technical details of the sampling step, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.

\subsection{Calibration Step}
\label{calibration_step}
Given that the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.  This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step. First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}. Given that for any $\beta \in \R / \{0\}$,
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
 &=&   \boldsymbol{M}_k \left(\boldsymbol{b} (1) \beta \right) \left(X_{t, k}(1)/\beta\right)  + \boldsymbol{v}_{t, k}\\
&=&   \boldsymbol{M}_k \boldsymbol{b} (\beta) X_{t, k}(\beta) + \boldsymbol{v}_{t, k},
\end{eqnarray*}
the parameter values under $b_{3} = \beta$ can be obtained from $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
%Furthermore, an estimate of $X_{t, k}(\beta)$ is given by $\hat{X}_{t, k}(1)/\beta$. 
This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$. Given that the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$. A natural criterion for this maximization is given by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}). Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions. Therefore,
\begin{eqnarray}
 \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right) \label{OSE}
\end{eqnarray}
where $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
\begin{eqnarray*}
S_{BRI}(Z, X) &=& -(Z - \logit^{-1}(X))^2
\end{eqnarray*}
or the logarithmic score (\citet{good1952rational})
\begin{eqnarray*}
S_{LOG}(Z, X) &=& Z \log\left(\logit^{-1}(X)\right) + (1-Z) \log\left(1-\logit^{-1}(X)\right)
\end{eqnarray*}
Given that it is not clear which rule should be used for predicting geopolitical events, the SAC-procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}. Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
\begin{eqnarray}
 \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \nonumber\\
 \hat{\boldsymbol{b}}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1)/ \hat{\beta}^2\nonumber\\
  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)\nonumber
\end{eqnarray}
Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint. Therefore their constrained and unconstrained versions are the same.

@@

Both $\text{SAC}_{\text{LOG}}$ and  $\text{SAC}_{\text{BRI}}$ estimate the hidden process and the bias vector very accurately. Apart from $\text{SAC}_{\text{LOG}}$ outperforming $\text{SAC}_{\text{BRI}}$ uniformly across all parameter values, the two approaches behave very similarly.  They suffer  from the increasing level of noise in the expert logit-probabilities, estimate the bias vector less accurately when $\beta$ is large, and, unlike EWMA, make greater use of the increasing number of questions.  EWMA is able to outperform SAC only when the bias among the experts or the number of problems is very small.

@@

\item  \textit{Exponentially Weighted Moving Logit Aggregator (EWMLA)}. This is a moving version of the aggregator $\hat{p}_G(\boldsymbol{b})$ that was introduced in \cite{satopaa}. If $\boldsymbol{p}_{t,k}$ is a vector collecting all the probability forecasts made for the $k$th question at time $t$, then the EWMLA forecasts are found recursively from
\begin{eqnarray*}
\hat{p}_{t,k}(\alpha, \boldsymbol{b}) =
\begin{cases}
G_{1,k}(\boldsymbol{b} ), & \text{ for } t  = 1, \\
\alpha G_{t,k}(\boldsymbol{b} )  + (1-\alpha) \hat{p}_{t-1,k}(\alpha, \boldsymbol{b}),  & \text{ for } t > 1,
\end{cases}
\end{eqnarray*}
where
\small
\begin{eqnarray*}
G_{t,k}(\nu )  &=& \left( \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{ \frac{\boldsymbol{e}_{i,k}'\boldsymbol{b}}{N_{t,k}}} \right) \Bigg/ \left(1+ \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{\frac{\boldsymbol{e}_{i,k}' \boldsymbol{b}}{N_{t,k}}} \right)
\end{eqnarray*}
\normalsize
The vector $\boldsymbol{b}$ collects the bias terms of the different expertise groups. Therefore it is equivalent to the bias vector found under \textit{Sample-And-Calibrate}. The term $\boldsymbol{e}_{i,k}$ is a vector of length 5 indicating which level of self-reported expertise the $i$th expert in the $k$th question belongs to. For instance, if $\boldsymbol{e}_{i,k} = [0, 1, 0, 0, 0]$, then the expert identifies himself with the expertise level two. The tuning parameters $(\alpha, \boldsymbol{b})$ are learned from the training set by
\begin{eqnarray*}
(\hat{\alpha}, \hat{ \boldsymbol{b}}) &=& \argmin_{ \boldsymbol{b} \in \R^5, \alpha \in [0,1]} \sum_{k \in S_{train}} \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}(\alpha,  \boldsymbol{b})\right)^2
\end{eqnarray*}

@@

Table \ref{prediction} summarizes different ways to aggregate these scores. The first option, denoted by \textit{Scores by Day}, weighs each question by the number of days the question remained open. This is performed by computing the average of the 17,475 scores. The second option, denoted by \textit{Scores by Problem}, gives each question an equal weight regardless how long the question remained open. This is done by first averaging the scores within a question and then averaging the average scores across all the questions. Both scores can be further broken down into subcategories by considering the length of the questions. The final three columns of Table \ref{prediction} divide the questions into \textit{Short} questions (30 days or fewer), \textit{Medium} questions (between 31 and 59 days), and \textit{Long} Problems (60 days or more). The number of questions in these subcategories were 36, 32 and 98, respectively. The bolded scores indicate the lowest score in each column. The values in the parenthesis quantify the variability in the scores: Under \textit{Scores by Day} the values give the standard errors of all the scores. Under \textit{Scores by Problem}, on other hand, the values represent the standard errors of the average scores of the different questions.

\begin{table*}
   \caption{Brier Scores based on 10-fold cross-validation. \textit{Scores by Day} weighs a question by the number of days the question remained open. \textit{Scores by Problem} gives each question an equal weight regardless how long the question remained open. The bolded values indicate the lowest scores in each column. The values in the parenthesis represent standard errors in the scores. }
   \label{prediction}
   \centering
      \begin{tabular}{l llll} % Column formatting, @{} suppresses leading/trailing space
      \hline
 \multicolumn{5}{ c }{Scores by Day}\\
Model & \multicolumn{1}{ c }{All} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
SDLM & 0.100 (0.156) & 0.066 (0.116) & 0.098 (0.154) & 0.102 (0.157)\\ 
$\text{BSAC}_{\text{LOG}}$ & 0.097 (0.213) & \textbf{0.053} (0.147) & 0.100 (0.215) & 0.098 (0.215)\\ 
$\text{SAC}_{\text{BRI}}$ & 0.096 (0.190) & 0.056 (0.134) & 0.097 (0.190) & 0.098 (0.192)\\ 
$\text{SAC}_{\text{LOG}}$ & \textbf{0.096} (0.191) & 0.056 (0.134) & \textbf{0.096} (0.189) & \textbf{0.098} (0.193)\\ 
EWMBA & 0.102 (0.203) & 0.060 (0.124) & 0.110 (0.201) & 0.103 (0.206)\\ 
EWMLA & 0.102 (0.199) & 0.061 (0.130) & 0.111 (0.214) & 0.103 (0.200)\\ 
EWMA & 0.111 (0.142) & 0.089 (0.100) & 0.111 (0.136) & 0.112 (0.144)\\ 
\hline
\multicolumn{5}{ c }{Scores by Problem}\\
Model & \multicolumn{1}{ c }{All} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
SDLM & 0.089 (0.116) & 0.064 (0.085) & 0.106 (0.141) & 0.092 (0.117)\\ 
$\text{BSAC}_{\text{LOG}}$ & 0.083 (0.160) & \textbf{0.052} (0.103) & 0.110 (0.198) & 0.085 (0.162)\\ 
$\text{SAC}_{\text{BRI}}$& 0.083 (0.142) & 0.055 (0.096) & 0.106 (0.174) & 0.085 (0.144)\\ 
$\text{SAC}_{\text{LOG}}$ & \textbf{0.082} (0.142) & 0.055 (0.096) & \textbf{0.105} (0.174) & \textbf{0.085} (0.144)\\ 
EWMBA & 0.090 (0.156) & 0.063 (0.101) & 0.118 (0.186) & 0.091 (0.161)\\ 
EWMLA & 0.090 (0.159) & 0.064 (0.109) & 0.120 (0.200) & 0.090 (0.159)\\ 
EWMA & 0.104 (0.105) & 0.092 (0.081) & 0.119 (0.125) & 0.103 (0.107)\\ 
\hline
   \end{tabular}
\end{table*}

Overall, $\text{SAC}_{\text{BRI}}$ and $\text{SAC}_{\text{LOG}}$ achieve the lowest average scores across all columns except \textit{Short}, where they are slightly outperformed by $\text{BSAC}_{\text{LOG}}$.  $\text{BSAC}_{\text{LOG}}$, however, turns out to be overconfident (see Section \ref{calibration}). This means that $\text{BSAC}_{\text{LOG}}$ underestimates the uncertainty in the events and outputs probability forecasts that are typically too near 0.0 or 1.0. As a result, the aggregate forecasts are either very close to the correct answer or very far from it.  As Table \ref{prediction} shows, this results into highly variable forecasting performance. The short questions, however, involve very little uncertainty and were generally the easiest to forecast. On such easy questions, overconfidence can pay off frequently enough to compensate for a few large losses arising from the overconfident and incorrect forecasts.

@@

 Similarly behavior is exhibited by EWMA that performs the worst of all the competing models. In the contrary, the other two exponentially weighted aggregators, EWMLA and EWMBA, make efficient use of the training set and present moderate forecasting performance in most columns of Table \ref{prediction}. Recall that EWMBA uses the cumulative distribution function of the Beta distribution that depends on two parameters and is more flexible than the transformation used by EWMLA. On other hand, only EWLMA is given access to the self-reported expertise information. Neither approach, however, appears to dominate the other. The aggregators perform very similarly. The high variability and average of their performance scores indicates that their performance suffers from over-confidence.

@@

\subsection{In- and Out-of-Sample Sharpness and Calibration}
\label{calibration}
A calibration plot is a simple tool for visually assessing the sharpness and calibration of a model. The idea is to plot the probability forecasts against the observed empirical frequencies. Therefore any deviation from the diagonal line suggests poor calibration. A model is considered under-confident (or over-confidence) if the points follow an S-shaped (or \reflectbox{S}-shaped) trend. To assess sharpness of the model, it is common practice to place a histogram of the given forecasts in the corner of the plot. Given that the data were balanced, any deviation from the the baseline probability of 0.5 suggests improved sharpness.

@@

The decreasing trend in under-confidence can be reasoned by viewing the process of making a subjective probability  as Bayesian updating: A completely ignorant expert aiming to minimize a reasonable loss function, such as the Brier score, has no reason to give anything but 0.5 as his probability forecast. However, as soon as the expert gains some knowledge about the event, he produces an updated forecast that is a compromise between his initial forecast and the new information acquired. The updated forecast is therefore conservative and too close to 0.5 as long as the expert remains only partially informed about the event. If most experts fall somewhere on this spectrum between ignorance and full information, their average forecast tends to fall strictly between 0.5 and the most-informed probability forecast (see \citet{Baron} for more details). Given that expertise is to a large extent determined by subject-matter knowledge, the level of under-confidence can be expected to decrease as a function of the group's level of self-reported expertise.

>>>>

\begin{aug}
\author{\fnms{Ville A.} \snm{Satop\"a\"a}\corref{}\ead[label=e1]{satopaa@wharton.upenn.edu}},
\author{\fnms{Shane T.} \snm{Jensen}\ead[label=e3]{stjensen@wharton.upenn.edu}},
\author{\fnms{Barbara A.} \snm{Mellers}\ead[label=e4]{mellers@wharton.upenn.edu}},
\author{\fnms{Philip E.} \snm{Tetlock}\ead[label=e5]{tetlock@wharton.upenn.edu}},
\and
\author{\fnms{Lyle H.} \snm{Ungar}\ead[label=e2]{ungar@cis.upenn.edu}}

@@

This paper begins by describing our geopolitical database. The paper then introduces a dynamic hierarchical model for capturing the crowd belief. The model is estimated in a two-step procedure: first, a sampling step produces constrained parameter estimates via Gibbs sampling (see, e.g., \citet{geman1984stochastic}); second, a calibration step transforms these estimates to their unconstrained equivalents via a one-dimension optimization procedure. The model introduction is followed by the first evaluation section that uses synthetic data to study how accurately the two-step procedure can estimate parameter values. The second evaluation section applies the model to our real-world geopolitical forecasting database. The paper concludes with a discussion on future research directions and model limitations.

@@

To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}. The points have been jittered slightly to make overlaps visible. The darkness of the points is positively correlated with the self-reported expertise. Given that the European bailout fund was ratified before November 1, 2011 and that the Nikkei 225 index finished trading at around 8,700 on September 30, 2011, the general trend of the probability forecasts tends to converge towards the correct answers. The individual experts, however, sometimes disagree strongly, with the disagreement persisting even near the closing dates of the questions.

@@

\subsection{Sampling Step}
\label{sampling_step}
Given that $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable. As a result, the parameter estimates tend to drift during the sampling process. A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.  In Section \ref{syntheticData} we provide a guideline for choosing the reference point. Denote the constrained version of the model by
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
X_{t, k}(1) &=& \gamma_k(1) X_{t-1, k}(1) + w_{t, k}\\
\boldsymbol{v}_{t, k} | \sigma^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k(1) \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}\left(0, \tau^2_k(1)\right),
\end{eqnarray*}
where the trailing input parameter emphasizes the constraint $b_{3} = 1$. Given that this version is identifiable, estimates of the model parameters can be obtained. Denote the estimates by placing a hat on the parameter symbol. For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively. These estimates are found via Gibbs sampling that only makes use of standard distributions. See Appendix \ref{appendix} for the technical details of the sampling steps, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.

\subsection{Calibration Step}
\label{calibration_step}
Given that the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.  This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step. First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}. Given that for any $\beta \in \R / \{0\}$,
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
 &=&   \boldsymbol{M}_k \left(\boldsymbol{b} (1) \beta \right) \left(X_{t, k}(1)/\beta\right)  + \boldsymbol{v}_{t, k}\\
&=&   \boldsymbol{M}_k \boldsymbol{b} (\beta) X_{t, k}(\beta) + \boldsymbol{v}_{t, k},
\end{eqnarray*}
the parameter values under $b_{3} = \beta$ can be obtained from $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
%Furthermore, an estimate of $X_{t, k}(\beta)$ is given by $\hat{X}_{t, k}(1)/\beta$. 
This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$. Given that the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$. A natural criterion for this maximization is given by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}). Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions. Therefore,
\begin{eqnarray}
 \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right) \label{OSE}
\end{eqnarray}
where $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
\begin{eqnarray*}
S_{BRI}(Z, X) &=& -(Z - \logit^{-1}(X))^2
\end{eqnarray*}
or the logarithmic score (\citet{good1952rational})
\begin{eqnarray*}
S_{LOG}(Z, X) &=& Z \log\left(\logit^{-1}(X)\right) + (1-Z) \log\left(1-\logit^{-1}(X)\right)
\end{eqnarray*}
Given that it is not clear which rule should be used for predicting geopolitical events, the SAC-procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}. Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
\begin{eqnarray}
 \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \nonumber\\
 \hat{\boldsymbol{b}}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1)/ \hat{\beta}^2\nonumber\\
  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)\nonumber
\end{eqnarray}
Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint.

@@

Both $\text{SAC}_{\text{LOG}}$ and  $\text{SAC}_{\text{BRI}}$ estimate the hidden process and the bias vector very accurately. Apart from $\text{SAC}_{\text{LOG}}$ outperforming $\text{SAC}_{\text{BRI}}$ uniformly across all parameter values, the two approaches behave very similarly.  They suffer  from the increasing level of noise in the expert logit-probabilities, estimate the bias vector less accurately when $\beta$ is large, and, unlike EWMA, make greater use of the increasing number of questions.  EWMA is able to outperform SAC only when the bias among the experts is very small.

@@

\item  \textit{Exponentially Weighted Moving Logit Aggregator (EWMLA)}. This is a moving version of the aggregator $\hat{p}_G(\boldsymbol{b})$ that was introduced in \cite{satopaa}. If $\boldsymbol{p}_{t,k}$ is a vector collecting all the probability forecasts made for the $k$th question at time $t$, then the EWMLA forecasts are found recursively from
\begin{eqnarray*}
\hat{p}_{t,k}(\alpha, \boldsymbol{b}) =
\begin{cases}
G_{1,k}(\boldsymbol{b} ), & \text{ for } t  = 1, \\
\alpha G_{t,k}(\boldsymbol{b} )  + (1-\alpha) \hat{p}_{t-1,k}(\alpha, \boldsymbol{b}),  & \text{ for } t > 1,
\end{cases}
\end{eqnarray*}
where
\small
\begin{eqnarray*}
G_{t,k}(\nu )  &=& \left( \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{ \frac{\boldsymbol{e}_{i,k}'\boldsymbol{b}}{N_{t,k}}} \right) \Bigg/ \left(1+ \prod\limits_{i=1}^{N_{t,k}} \left( \frac{p_{i, t, k}}{1-p_{i, t, k}} \right)^{\frac{\boldsymbol{e}_{i,k}' \boldsymbol{b}}{N_{t,k}}} \right)
\end{eqnarray*}
\normalsize
The vector $\boldsymbol{b}$ collects the bias terms of the different expertise groups. Therefore it serves a similar role to the bias vector found under \textit{Sample-And-Calibrate}. The term $\boldsymbol{e}_{i,k}$ is a vector of length 5 indicating which level of self-reported expertise the $i$th expert in the $k$th question belongs to. For instance, if $\boldsymbol{e}_{i,k} = [0, 1, 0, 0, 0]$, then the expert identifies himself with the expertise level two. The tuning parameters $(\alpha, \boldsymbol{b})$ are learned from the training set by
\begin{eqnarray*}
(\hat{\alpha}, \hat{ \boldsymbol{b}}) &=& \argmin_{ \boldsymbol{b} \in \R^5, \alpha \in [0,1]} \sum_{k \in S_{train}} \sum_{t=1}^{T_k} \left(Z_k - \hat{p}_{t,k}(\alpha,  \boldsymbol{b})\right)^2
\end{eqnarray*}

@@

Table \ref{prediction} summarizes different ways to aggregate these scores. The first option, denoted by \textit{Scores by Day}, weighs each question by the number of days the question remained open. This is performed by computing the average of the 17,475 scores. The second option, denoted by \textit{Scores by Problem}, gives each question an equal weight regardless how long the question remained open. This is done by first averaging the scores within a question and then averaging the average scores across all the questions. Both scores can be further broken down into subcategories by considering the length of the questions. The final three columns of Table \ref{prediction} divide the questions into \textit{Short} questions (30 days or fewer), \textit{Medium} questions (between 31 and 59 days), and \textit{Long} Problems (60 days or more). The number of questions in these subcategories were 36, 32 and 98, respectively. The bolded scores indicate the best score in each column. The values in the parenthesis quantify the variability in the scores: Under \textit{Scores by Day} the values give the standard errors of all the scores. Under \textit{Scores by Problem}, on other hand, the values represent the standard errors of the average scores of the different questions.

\begin{table*}
   \caption{Brier Scores based on 10-fold cross-validation. \textit{Scores by Day} weighs a question by the number of days the question remained open. \textit{Scores by Problem} gives each question an equal weight regardless how long the question remained open. The bolded values indicate the best scores in each column. The values in the parenthesis represent standard errors in the scores. }
   \label{prediction}
   \centering
      \begin{tabular}{l llll} % Column formatting, @{} suppresses leading/trailing space
      \hline
 \multicolumn{5}{ c }{Scores by Day}\\
Model & \multicolumn{1}{ c }{All} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
SDLM & 0.100 (0.156) & 0.066 (0.116) & 0.098 (0.154) & 0.102 (0.157)\\ 
$\text{BSAC}_{\text{LOG}}$ & 0.097 (0.213) & \textbf{0.053} (0.147) & 0.100 (0.215) & 0.098 (0.215)\\ 
$\text{SAC}_{\text{BRI}}$ & 0.096 (0.190) & 0.056 (0.134) & 0.097 (0.190) & 0.098 (0.192)\\ 
$\text{SAC}_{\text{LOG}}$ & \textbf{0.096} (0.191) & 0.056 (0.134) & \textbf{0.096} (0.189) & \textbf{0.098} (0.193)\\ 
EWMBA & 0.102 (0.203) & 0.060 (0.124) & 0.110 (0.201) & 0.103 (0.206)\\ 
EWMLA & 0.102 (0.199) & 0.061 (0.130) & 0.111 (0.214) & 0.103 (0.200)\\ 
EWMA & 0.111 (0.142) & 0.089 (0.100) & 0.111 (0.136) & 0.112 (0.144)\\ 
\hline
\multicolumn{5}{ c }{Scores by Problem}\\
Model & \multicolumn{1}{ c }{All} & \multicolumn{1}{ c }{Short} & \multicolumn{1}{ c }{Medium} & \multicolumn{1}{ c }{Long}\\ \hline
SDLM & 0.089 (0.116) & 0.064 (0.085) & 0.106 (0.141) & 0.092 (0.117)\\ 
$\text{BSAC}_{\text{LOG}}$ & 0.083 (0.160) & \textbf{0.052} (0.103) & 0.110 (0.198) & 0.085 (0.162)\\ 
$\text{SAC}_{\text{BRI}}$& 0.083 (0.142) & 0.055 (0.096) & 0.106 (0.174) & 0.085 (0.144)\\ 
$\text{SAC}_{\text{LOG}}$ & \textbf{0.082} (0.142) & 0.055 (0.096) & \textbf{0.105} (0.174) & \textbf{0.085} (0.144)\\ 
EWMBA & 0.090 (0.156) & 0.063 (0.101) & 0.118 (0.186) & 0.091 (0.161)\\ 
EWMLA & 0.090 (0.159) & 0.064 (0.109) & 0.120 (0.200) & 0.090 (0.159)\\ 
EWMA & 0.104 (0.105) & 0.092 (0.081) & 0.119 (0.125) & 0.103 (0.107)\\ 
\hline
   \end{tabular}
\end{table*}

Overall, $\text{SAC}_{\text{BRI}}$ and $\text{SAC}_{\text{LOG}}$ achieve the lowest average scores across all columns except \textit{Short}, where they are slightly outperformed by $\text{BSAC}_{\text{LOG}}$.  However, $\text{BSAC}_{\text{LOG}}$ turns out to be overconfident (see Section \ref{calibration}). This means that $\text{BSAC}_{\text{LOG}}$ underestimates the uncertainty in the events and outputs probability forecasts that are typically too near 0.0 or 1.0. As a result, the aggregate forecasts are either very close to the correct answer or very far from it.  As Table \ref{prediction} shows, this results into highly variable forecasting performance. The short questions, however, involve very little uncertainty and were generally the easiest to forecast. On such easy questions, overconfidence can pay off frequently enough to compensate for a few large losses arising from the overconfident and incorrect forecasts.

@@

 Similar behavior is exhibited by EWMA that performs the worst of all the competing models. In contrast, the other two exponentially weighted aggregators, EWMLA and EWMBA, make efficient use of the training set and present moderate forecasting performance in most columns of Table \ref{prediction}. Recall that EWMBA uses the cumulative distribution function of the Beta distribution that depends on two parameters and is more flexible than the transformation used by EWMLA. On other hand, only EWLMA is given access to the self-reported expertise information. Neither approach, however, appears to dominate the other. The aggregators perform very similarly. The high variability and average of their performance scores indicate that their performance suffers from over-confidence.

@@

\subsection{In- and Out-of-Sample Sharpness and Calibration}
\label{calibration}
A calibration plot is a simple tool for visually assessing the sharpness and calibration of a model. The idea is to plot the probability forecasts against the observed empirical frequencies. Therefore any deviation from the diagonal line suggests poor calibration. A model is considered under-confident (or over-confident) if the points follow an S-shaped (or \reflectbox{S}-shaped) trend. To assess sharpness of the model, it is common practice to place a histogram of the given forecasts in the corner of the plot. Given that the data were balanced, any deviation from the the baseline probability of 0.5 suggests improved sharpness.

@@

The decreasing trend in under-confidence can be reasoned by viewing the process of making a subjective probability  as Bayesian updating: a completely ignorant expert aiming to minimize a reasonable loss function, such as the Brier score, has no reason to give anything but 0.5 as his probability forecast. However, as soon as the expert gains some knowledge about the event, he produces an updated forecast that is a compromise between his initial forecast and the new information acquired. The updated forecast is therefore conservative and too close to 0.5 as long as the expert remains only partially informed about the event. If most experts fall somewhere on this spectrum between ignorance and full information, their average forecast tends to fall strictly between 0.5 and the most-informed probability forecast (see \citet{Baron} for more details). Given that expertise is to a large extent determined by subject-matter knowledge, the level of under-confidence can be expected to decrease as a function of the group's level of self-reported expertise.
